Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-07-06 write-ups/FSE/paper.tex
Done.

\myparagraph{Responsiveness}
%
\myfiglong\ref{fig:fine-grained-timing} reports, on three
configurations, the average gather and execution times for the
non-cooperative kernel w.r.t. the quantity of workgroups allocated to
it.  A logarithmic scale is used for time since gather times tend to
be much smaller than execution times. The horizontal grey lines
indicates the desired period for non-cooperative kernels.  These
graphs show a representative sample of our results; the full set of
graphs for all configurations is provided in supplementary material.

The left most graph illustrates a work
stealing example.  When the non-cooperative kernel is given only one
workgroup, its execution is so long that it cannot complete within the
period required for a screen refresh. The gather time is very good
though, since the scheduler needs to collect only one workgroup.  The
more workgroups are allocated to the non-cooperative kernels, the
faster it can compute: here the non-cooperative kernel becomes fast
enough with a quarter (resp.\ half) of available workgroups for light
(resp.\ heavy) graphics workload. Inversely, the gather time increases
since the scheduler must collect more and more workgroups.

The middle and right graphs show results for graph algorithms.  These
algorithms use barriers, and we experimented with the regular and
query barrier implementations described in
\mysec\ref{sec:resizingbarrier}.  The execution times for the
non-cooperative task are averaged across all runs, including with both
types of barrier.  We show separately the average gather time
associated with each type of barrier.  The graphs show a similar trend
to the left-most graph: as the number of non-cooperative workgroups
grows, the execution time decreases and the gather time
increases.%\footnote{The right figure has only 3 points per line:
  %recall that for bfs there are no runs with ``all-but-one''
  %non-cooperative workgroups.}
%
the gather time is higher on the right figure: the G3 circuit input
graph is rather wide than deep, so the graph algorithm reaches
resizing barriers less often than for the USA road input of the middle
figure for instance. The scheduler thus has fewer opportunities to
collect workgroups and gather time increases. Nonetheless, scheduling
responsiveness can benefit from the query barrier: when used, this
barrier lets the scheduler collect all needed workgroups as soon as
they hit a resizing barrier. 
As we can see, the gather time of the
query barrier is almost stable w.r.t.\ the number of workgroups that
needs to be collected.

@@

\myparagraph{Performance} \myfiglong\ref{fig:performance} reports the
overhead brought by the scheduling of non-cooperative kernels over the
cooperative kernel execution time.  This is the slowdown associated
with running the cooperative kernel in the presence of multitasking,
vs.\ running the cooperative kernel in isolation (median over all
applications and inputs).  We also show the period at which
non-cooperative kernels can be scheduled (median over all applications
and inputs). Our data included some outliers that occur with
benchmarks in which the resizing barrier are not called very
frequently and the graphics task requires half or more workgroups. For
example, a medium graphics workload for bfs on the rmat input has over
an 8$\times$ overhead when asking for all but one of the
workgroups. As \myfiglong\ref{fig:performance} shows, most of our
benchmarks are much better behaved than this. Future work
will examine these problem benchmarks in more detail,
possibly inserting more resizing calls.

@@

Co-scheduling non-cooperative kernels that request a single workgroup
leads to almost no overhead, but the period is far too high to meet
the needs of any of our three workloads; e.g.\ a heavy workload
averages a period of 939 $\mathit{ms}$. As more workgroups are dedicated to
non-cooperative kernels, they execute quickly enough to be scheduled
at the expected period. For the light and medium workloads, a quarter
of the workgroups executing the non-cooperative kernel are able to
meet their goal period (70 and 40 $\mathit{ms}$ resp.). However, this is not
sufficient to meet the goal for the heavy workload (giving a median
period of 104 $\mathit{ms}$). If half of the workgroups are allocated to the
non-cooperative kernel, the heavy workload achieves its goal period
(median of 40 $\mathit{ms}$).
%
Yet, as expected, allocating more non-cooperative workgroups increases
the overhead of the cooperative kernel.
%
Still, heavy workloads meet their period by allocating half
of the workgroups, incurring a slow down of less than
1.5$\times$ (median). Light and medium workloads meet their period
with only a small overhead; 1.04$\times$ and 1.08$\times$ median
slowdown respectively.

>>>>

\myparagraph{Responsiveness}
%
\myfiglong\ref{fig:fine-grained-timing} reports, on three
configurations, the average gather and execution times for the
non-cooperative kernel with respect to the quantity of workgroups allocated to
it.  A logarithmic scale is used for time since gather times tend to
be much smaller than execution times. The horizontal grey lines
indicates the desired period for non-cooperative kernels.  These
graphs show a representative sample of our results; the full set of
graphs for all configurations is provided in supplementary material.

The left-most graph illustrates a work
stealing example.  When the non-cooperative kernel is given only one
workgroup, its execution is so long that it cannot complete within the
period required for a screen refresh. The gather time is very good
though, since the scheduler needs to collect only one workgroup.  The
more workgroups are allocated to the non-cooperative kernels, the
faster it can compute: here the non-cooperative kernel becomes fast
enough with a quarter (resp.\ half) of available workgroups for light
(resp.\ heavy) graphics workload. Inversely, the gather time increases
since the scheduler must collect more and more workgroups.

The middle and right graphs show results for graph algorithms.  These
algorithms use barriers, and we experimented with the regular and
query barrier implementations described in
\mysec\ref{sec:resizingbarrier}.  The execution times for the
non-cooperative task are averaged across all runs, including with both
types of barrier.  We show separately the average gather time
associated with each type of barrier.  The graphs show a similar trend
to the left-most graph: as the number of non-cooperative workgroups
grows, the execution time decreases and the gather time
increases.%\footnote{The right figure has only 3 points per line:
  %recall that for bfs there are no runs with ``all-but-one''
  %non-cooperative workgroups.}
%
the gather time is higher on the right figure: the G3 circuit input
graph is rather wide than deep, so the graph algorithm reaches
resizing barriers less often than for the USA road input of the middle
figure for instance. The scheduler thus has fewer opportunities to
collect workgroups and gather time increases. Nonetheless, scheduling
responsiveness can benefit from the query barrier: when used, this
barrier lets the scheduler collect all needed workgroups as soon as
they hit a resizing barrier. 
As we can see, the gather time of the
query barrier is almost stable with respect to the number of workgroups that
needs to be collected.

@@

\myparagraph{Performance} \myfiglong\ref{fig:performance} reports the
overhead brought by the scheduling of non-cooperative kernels over the
cooperative kernel execution time.  This is the slowdown associated
with running the cooperative kernel in the presence of multitasking,
vs.\ running the cooperative kernel in isolation (median over all
applications and inputs).  We also show the period at which
non-cooperative kernels can be scheduled (median over all applications
and inputs). Our data included some outliers that occur with
benchmarks in which the resizing barrier are not called very
frequently and the graphics task requires half or more workgroups. For
example, a medium graphics workload for bfs on the rmat input has over
an 8$\times$ overhead when asking for all but one of the
workgroups. As \myfiglong\ref{fig:performance} shows, most of our
benchmarks are much better behaved than this. In future work
is required to examine the problematic benchmarks in more detail,
possibly inserting more resizing calls.

@@

Co-scheduling non-cooperative kernels that request a single workgroup
leads to almost no overhead, but the period is far too high to meet
the needs of any of our three workloads; e.g.\ a heavy workload
averages a period of $939\mathit{ms}$. As more workgroups are dedicated to
non-cooperative kernels, they execute quickly enough to be scheduled
at the expected period. For the light and medium workloads, a quarter
of the workgroups executing the non-cooperative kernel are able to
meet their goal period (70 and $40\mathit{ms}$ resp.). However, this is not
sufficient to meet the goal for the heavy workload (giving a median
period of $104\mathit{ms}$). If half of the workgroups are allocated to the
non-cooperative kernel, the heavy workload achieves its goal period
(median of $40\mathit{ms}$).
%
Yet, as expected, allocating more non-cooperative workgroups increases
the overhead of the cooperative kernel.
%
Still, heavy workloads meet their period by allocating half
of the workgroups, incurring a slow down of less than
1.5$\times$ (median). Light and medium workloads meet their period
with only a small overhead; 1.04$\times$ and 1.08$\times$ median
slowdown respectively.
